Which ITG Stepcharts are Turniest?

### 2019-03-17 itg/3/paper.tex
proofread

What, then, is a bracket-jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
(see \cite{turniness}),
while a novice player might think they must step with both feet at once, one for each arrow,
experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
to hit both arrows by
triggering one arrow's sensor with the heel and the other with the little bitty toesies.
Pads are typically constructed with a small triangular metal bracket at the corner of each arrow panel,
as shown in \Cref{fig:bracker-detail},
which the bridge of the foot must cross to achieve this, hence the name ``bracket jump''.
In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
a high-quality graphics render of a player's typical foot positions during a down+right bracket-jump (henceforth ``DR'', et cetera).

@@

I extended the crossoveriness et cetera algorithm from \cite{crossoveriness} to reason about jumps,
which previously it treated all identically,
ignoring the arrows involved and allowing the player to reset her footing as desired.
Now, it considers LD, LU, DR, and UR jumps as potentially bracketable,
allowing the player to continue a stream of alternating feet uninterrupted through.

@@

The general gist is that in addition to tracking which foot stepped the last arrow and optimizing for alternating feet,
we also track which arrow(s) each foot was last on,
and whenever we encounter a LD/LU/DR/UR jump, bracket it if both:
\begin{enumerate}
	\item The last foot is opposite the foot required for the jump (e.g. L for LD), and
	\item The last foot's last note(s) does not intersect the arrows involved in the jump (i.e., R not on D preceding LD).
\end{enumerate}
In lieu of a detailed algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how the implementation must handle them,
which was pretty much my implementation strategy to begin with.

@@

\renewcommand{\theenumi}{\alph{enumi}}
\begin{enumerate}
	\item In the OK case, the right foot can bracket every jump (condition 1),
		as the left foot is always on a different arrow than the jump (condition 2).
		In the NG case, condition 2 is violated for the second and fourth jumps.
	\item Version of (a) testing that preceding bracket-jumps also contribute to condition 2.
		The NG case would be a footswitch bracket requiring rather uncomfortable footing angles.
	\item In OK, LR cannot be bracketed, but clears the prior footings so UR can.
		In NG, LR can be bracketed, meaning UR cannot while still alternating feet (condition 1).
	\item DU jumps can never be bracketed, but can be stepped facing either direction, and this must respect prior footings.
		In both cases DU is stepped with right foot U, which allows bracketing LD but not LU thereafter.
	\item In OK, the crossover does not interfere with either bracket and the player can alternate feet throughout.
		In NG, either doublestepping or jumping normally is required
		(or a future-work crossover bracket; see \Cref{sec:xover-brackers}).
	\item In OK, the player easily footswitches on U.
		In the corresponding case, the player can actually just jack U instead, stepping it twice with her left foot
		before bracketing DR again with her right.
\end{enumerate}
\renewcommand{\theenumi}{\arabic{enumi}}

@@

Next up is bracket jumpiness by pack and by author.
\Cref{tab:pack-bjs,tab:pack-bjj} show the top 10 packs in both bracketiness metrics,
and \Cref{tab:author-bjs,tab:author-bjj} the top 10 authors
(filtered by having written at least 10 charts).
For the packs, I also list the release year of each;
note how despite the relatively even spread of years in BJS\%,
2018 dominates the BJJ\% leaderboard,
suggesting that older charts' bracket-jumps arose by chance simply from having a lot of jumps to begin with.
I'll come back to this point in \Cref{sec:eval-years}.
In honorable mentionth place is chart author Halogen-, whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
What a purist!

@@

I sought to prove \cite{dril}'s claim that bracket-jumps have skyrocketed in popularity in the most recent years of stepchart-making,
%I decided to do this
by finding the progression of bracketiness across past years of stepchart packs.
However,
actually assigning a firm release date to every pack on my hard drive turned out to be a feat of internet archaeology unto itself,
even involving archive.org for one step.
I'll spare you the details, but you can peruse them in the second spreadsheet of the dataset linked at the start of this section.
Ultimately, all packs older than 2013 had to be grouped together,
as no pack collection website, facebook group, or laptop filesystem could accurately date enough packs
before then to give meaningful sample sizes for each year.
I also decided to exclude tournament packs such as ECFA from this analysis,
which curate existing stepcharts written possibly long ago.

\textit{Statistical significance.}
Stepcharts thus dated, I then summed the total steps, jumps, and brackets published in each year,
and analyzed the resulting overall BJS\% and BJJ\% as a linear regression.
\Cref{fig:stats} shows the best-fit and 95\% confidence intervals,
plotted in R \cite{r-lang} with ggplot \cite{ggplot}.
For BJS\%, the $\le$2012 bucket actually produced the highest data point (at 11.8\%);
I suspect this simply represents the fact that jumps were more popular overall during ITG's nascency
(as corroborated by \Cref{tab:pack-bjs}).
Owing to this and also to its nature as an aggregation over nearly a decade (which spoils the linear model anyway),
I chose to exclude it from BJS\%,
but keep it for BJJ\%, which overall jump count has no bearing on.
(Its linear fit has $\beta$ = 0.012\%, CI = \{-0.08,+0.10\}, not significant.)
For the two pictured distributions, their confidence intervals do not include 0,
i.e., a flat line representing no growth,
so I conclude bracket jumps' increased popularity is statistically significant.

>>>>

What, then, is a bracket-jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
\cite{turniness},
while a novice player might think they must step with both feet at once, one for each arrow,
experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
to hit both arrows by
triggering one arrow's sensor with the heel and the other with the little bitty toesies.
Pads are typically constructed with a small triangular metal bracket at the corner of each arrow panel,
as shown in \Cref{fig:bracker-detail},
which the bridge of the foot must cross to achieve this, hence the name ``bracket jump''.
In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
a high-quality graphics render of a player's typical foot positions during a down+right bracket-jump (henceforth ``DR'', et cetera).

@@

I extended the crossoveriness et cetera algorithm from \cite{crossoveriness} to reason about jumps,
which previously it treated all identically,
ignoring the arrows involved and allowing the player to reset her footing as desired.
Now, it considers LD, LU, DR, and UR jumps as potentially bracketable,
allowing the player to continue a stream of alternating feet uninterrupted therethrough.

@@

The general gist is that in addition to tracking which foot stepped the last arrow and optimizing for alternating feet,
we also track which arrow(s) each foot was last on,
and whenever we encounter a LD/LU/DR/UR jump, bracket it if both:
\begin{enumerate}
	\item The last foot is opposite the foot required for the jump (e.g. L for LD), and
	\item The last foot's last note(s) does not intersect the arrows involved in the jump (i.e., R not on D preceding LD).
\end{enumerate}
In lieu of a detailed algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how a player would typically step them,
which was pretty much my implementation strategy to begin with.

@@

\renewcommand{\theenumi}{\alph{enumi}}
\begin{enumerate}
	\item In the OK case, the right foot can bracket every jump (condition 1),
		as the left foot is always on a different arrow than the jump (condition 2).
		In the NG case, condition 2 is violated for the second and fourth jumps.
	\item Version of (a) testing that preceding bracket-jumps also contribute to condition 2.
		The NG case would be a footswitch bracket requiring rather uncomfortable footing angles.
	\item In OK, DR cannot be bracketed, but clears the prior footings so UR can.
		In NG, LR can be bracketed, meaning UR cannot while still alternating feet (condition 1).
	\item DU jumps can never be bracketed, but can be stepped facing either direction, and this must respect prior footings.
		In both cases DU is stepped with right foot U, which allows bracketing LD but not LU thereafter.
	\item In OK, the crossover does not interfere with either bracket and the player can alternate feet throughout.
		In NG, either doublestepping or jumping normally is required
		(or a future-work crossover bracket; see \Cref{sec:xover-brackers}).
	\item In OK, the player easily footswitches on U.
		In the corresponding case, the player can actually just jack U instead, stepping it twice with her left foot
		before bracketing DR again with her right.
\end{enumerate}
\renewcommand{\theenumi}{\arabic{enumi}}

@@

Next up is bracket jumpiness by pack and by author.
\Cref{tab:pack-bjs,tab:pack-bjj} show the top 10 packs in both bracketiness metrics,
and \Cref{tab:author-bjs,tab:author-bjj} the top 10 authors
(filtered by having written at least 10 charts).
For the packs, I also list the release year of each;
note how despite the relatively even spread of years in BJS\%,
2018 dominates the BJJ\% leaderboard,
suggesting that older charts' bracket-jumps arose by chance simply from having a lot of jumps to begin with.
I'll come back to this point in \Cref{sec:eval-years}.
In honorable mentionth place is chart author \mbox{Halogen--,} whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
What a purist!

@@

I sought to prove \cite{dril}'s claim that bracket-jumps' popularity has skyrocketed in recent years of stepchart-making,
%I decided to do this
by finding the progression of bracketiness across past years of stepchart packs.
However,
actually assigning a firm release date to every pack on my hard drive turned out to be a feat of internet archaeology unto itself,
even involving archive.org for one step.
I'll spare you the details, but you can peruse them in the second spreadsheet of the dataset linked at the start of this section.
Ultimately, all packs older than 2013 had to be grouped together,
as no pack collection website, facebook group, or laptop filesystem metadata could accurately date enough packs
before then to give meaningful sample sizes for each year.
I also decided to exclude tournament packs such as ECFA from this analysis,
which curate existing stepcharts written possibly long ago.

\textit{Statistical significance.}
Stepcharts thus dated, I then summed the total steps, jumps, and brackets published in each year,
and analyzed the resulting overall BJS\% and BJJ\% as a linear regression.
\Cref{fig:stats} shows the best-fit and 95\% confidence intervals,
plotted in R \cite{r-lang} with ggplot \cite{ggplot}.
For BJS\%, the $\le$2012 bucket actually produced the highest data point (at 11.8\%);
I suspect this simply represents the fact that jumps were more popular overall during ITG's nascency
(as corroborated by \Cref{tab:pack-bjs}).
Owing to this and also to its nature as an aggregation over nearly a decade (which spoils the linear model anyway),
I chose to exclude it from BJS\%,
but kept it for BJJ\%, which overall jump count has no bearing on.
(Its linear fit has $\beta$ = 0.012\%, CI = \{-0.08,+0.10\}, not significant.)
For the two pictured distributions, their confidence intervals do not include 0,
i.e., a flat line representing no growth,
so I conclude bracket jumps' increased popularity is statistically significant.
